---
title: "hw8"
author: "Kunle Lawal, Anubhav Rana, Mihir Tulpule and Ali Lakdawala"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Stepwise models 
```{r}
require(data.table)
crimedata = read.table('uscrime.txt', header = TRUE)
library(MASS)

#Original Linear Model with all variables 
model <- lm(Crime ~., data = crimedata)
summary(model)

#Model with only significant variables 
model2 <- lm(Crime ~ Ed + Ineq + M +Po2 + Prob + U2, data = crimedata )
summary(model2)


#Stepwise with all
step_model <- step(model, direction = "both", trace = FALSE)
summary(step_model)

#Stepwise with only significant 
step_model2 <- step(model2, direction = "both", trace = FALSE)
summary(step_model2)





```
# Lasso Model & Elastic Net


```{r message = FALSE}
library(glmnet)
set.seed(1234)
scaleData <- scale(crimedata, center = TRUE, scale = TRUE)

x <- scaleData[,-16]
y <- scaleData[,16]

m <- nrow(x)
trn <- sample(1:m, size = round(m*0.7), replace = FALSE)
trainx <- x[trn,]
testx <- x[-trn,]


train_y <- y[trn]
test_y <- y[-trn]


fit.lasso <- glmnet(trainx, train_y, family = "gaussian", alpha = 1)
fit.elnet <- glmnet(trainx, train_y, family = "gaussian", alpha = 0.5)

R2 <- c()
for (i in 0:10) {
                                           
  elasticfit <- cv.glmnet(trainx, train_y, type.measure="mse", 
                          alpha=i/10,family="gaussian")
  R2 = cbind(R2, elasticfit$glmnet.fit$dev.ratio[which(elasticfit$glmnet.fit$lambda == elasticfit$lambda.min)])
}
R2

alphas <- (0:10)/10

plot(alphas, R2)
lines(alphas, R2)
```


In our original model with all the predictors  adjusted the Multiple R-Squared is 0.8031
In original model with only significant predictors have the Multiple R-Squared with 0.744

For Stepwise model with all predictors we noticed R Squared Values of 0.789
For Stepwise model with just significant we notice the same R squared as the original model with significant predictors 0.744.


Then for Elastic net and Lasso models we plotted the r-squared values above for alphas 0 -1. The maximum R-Squared value for the Elastic models was higher than the maximum value for the Lasso model but only by a very small margin. The maximum value is around 0.899. 

What we found is that Elastic and Lasso models were superior to the Stepwise model in terms of quality. 
We also notice that Stepwise regression and the orginal regression model with only significant predictors the R-squared value is the same. This makese sense as stepwise model chooses the top predictors and in this case they were all significant. 

It is important to note that all three methods produced models that were superior to the original linear regression model in terms of r-squared values or complexity. 




